Let's talk about Thurstone \& Co.: An information-theoretical model for comparative judgments, and its statistical translation

### 2025-04-11 _tex/paper2.tex
Built site for gh-pages

\section{From SCMs to a statistical model}\label{sec-statistical}

With the structure in SCM~\ref{fig-cj15_scm}, we can derive a
statistical model that addresses violations of the equal dispersion
assumption in CJ experiments (see Section~\ref{sec-theory-issue1a}).
This derivation is possible because a fully specified SCM encodes
functional and probabilistic information, which we can replace with
appropriate functions and probabilistic assumptions
\citep{Pearl_et_al_2016}. Specifically, SCM~\ref{fig-cj15_scm} allows us
to express the joint distribution of our CJ system as a product of
simpler conditional probability distributions (CPDs)\footnote{This
  re-expression is possible because the \emph{chain rule} of probability
  and the \emph{Bayesian Network Factorization (BNF)} property. For
  further details, see \citet{Pearl_et_al_2016} and \citet{Neal_2020}.},
as shown in equation (\ref{eq-mat51}). For clarity, we treat expressions
such as \(Y := f_{Y}(X)\), \(P(Y \mid X)\), and \(Y \sim f(Y \mid X)\)
as equivalent. Here, \(P(Y \mid X)\) and \(f(Y \mid X)\) represent the
CPD of \(Y\) given \(X\).
\begin{equation}\phantomsection\label{eq-mat51}{
\begin{aligned}
  P(O_{R}, & S, C, D_{R}, T_{IA}, X_{IA}, e_{IA}, T_{I}, X_{I}, e_{I}, B_{JK}, Z_{JK}, e_{JK}, B_{J}, Z_{J}, e_{J} ) & \\
  &= P(O_{R} \mid D_{R}, S, C) \cdot P(S) \cdot P(C) \cdot P(D_{R} \mid T_{IA}, B_{JK}) \\
  & \quad \cdot P(T_{IA} \mid T_{I}, X_{IA}, e_{IA}) \cdot P(T_{I} \mid X_{I}, e_{I}) \\
  & \quad \cdot P(B_{JK} \mid B_{J}, Z_{JK}, e_{JK}) \cdot P(B_{J} \mid Z_{J}, e_{J}) \\
  & \quad \cdot P(X_{IA}) \cdot P(X_{I}) \cdot P(Z_{JK}) \cdot P(Z_{J}) \\
  & \quad \cdot P(e_{IA}) \cdot P(e_{I}) \cdot P(e_{JK}) \cdot P(e_{J})  
\end{aligned}
}\end{equation}

>>>>

\section{From SCM to a statistical model}\label{sec-statistical}

With the structure in SCM~\ref{fig-cj15_scm}, we can derive a
statistical model that addresses violations of the equal dispersion
assumption in CJ experiments (see Section~\ref{sec-theory-issue1a}).
This derivation is possible because a fully specified SCM encodes
functional and probabilistic information, which we can replace with
appropriate functions and probabilistic assumptions
\citep{Pearl_et_al_2016}. Specifically, SCM~\ref{fig-cj15_scm} allows us
to express the joint distribution of our complex CJ system as a product
of simpler conditional probability distributions (CPDs)\footnote{This
  re-expression is possible because the \emph{chain rule} of probability
  and the \emph{Bayesian Network Factorization (BNF)} property. For
  further details, see \citet{Pearl_et_al_2016} and \citet{Neal_2020}.},
as shown in equation (\ref{eq-mat51}). For clarity, we treat expressions
such as \(Y := f_{Y}(X)\), \(P(Y \mid X)\), and \(Y \sim f(Y \mid X)\)
as equivalent. Here, \(P(Y \mid X)\) and \(f(Y \mid X)\) represent the
CPD of \(Y\) given \(X\).
\begin{equation}\phantomsection\label{eq-mat51}{
\begin{aligned}
  P(O_{R}, & S, C, D_{R}, T_{IA}, X_{IA}, e_{IA}, T_{I}, X_{I}, e_{I}, B_{JK}, Z_{JK}, e_{JK}, B_{J}, Z_{J}, e_{J} ) & \\
  &= P(O_{R} \mid D_{R}, S, C) \cdot P(S) \cdot P(C) \cdot P(D_{R} \mid T_{IA}, B_{JK}) \\
  & \quad \cdot P(T_{IA} \mid T_{I}, X_{IA}, e_{IA}) \cdot P(T_{I} \mid X_{I}, e_{I}) \\
  & \quad \cdot P(B_{JK} \mid B_{J}, Z_{JK}, e_{JK}) \cdot P(B_{J} \mid Z_{J}, e_{J}) \\
  & \quad \cdot P(X_{IA}) \cdot P(X_{I}) \cdot P(Z_{JK}) \cdot P(Z_{J}) \\
  & \quad \cdot P(e_{IA}) \cdot P(e_{I}) \cdot P(e_{JK}) \cdot P(e_{J})  
\end{aligned}
}\end{equation}
